---
title: "gentest -- R package for generative property-based testing with automatic doc generation"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(width=200)
library(magrittr)
genHTML <- function(filename) {
    readLines(filename) %T>%
        gsub('\\tab \\tab \\tab','\\tab',.,fixed=TRUE) %>%
        gsub('\\tab \\tab \\tab','\\tab',.,fixed=TRUE) %>%
        gsub('\\tab \\tab',"",.,fixed=TRUE) %>%
        cat(file=filename, sep='\n')
    capture.output(filename %>% tools::Rd2HTML(stdout())) %>%
        sub('<link rel="stylesheet" type="text/css" href="R.css" />',"",.,fixed=TRUE) %>%
        c('\\ \n', 'See the help file below (as rendered by GitHub):\n',
          '- - -', ., '- - -', '\\ \n') %>%
        cat(sep='\n')
    invisible(file.remove(filename))
}
```

## # API

#### ## Testing

- `gentest` -- evaluates a given function for all the combinations of
the arguments provided as vectors/lists; the results are presented
in a data frame (including possible errors and warnings)

#### ## Auto-Documentation

- `docgen` -- automatically generates the .Rd doc/help file from
the `gentest`'s result; it helps keeping your documentation
up-to-date and in sync with the test results
- `Name` -- use it to tag the return values of your own generator function
in a human-readable way, useful when `docgen` is used

#### ## Built-in generators of atomic vectors

- `Bool`
- `SingleCharacter`
- `String`
- `NumericInRange`
- `IntegerInRange`

#### ## Modifiers

- `Scalar` -- randomly selects a random element from a vector
(i.e. making the length of a vector == 1).
- `withNA` -- injects a fraction of `NA`s into a vector
- `withZero` -- injects a fraction of zeros (0) into a vector
(ensuring that zeros are included in the tests)

#### ## Wrappers

- `List` -- a wrapper for lists
- `DataFrame` -- a wrapper for data frames
- `%|%` -- a binary operator to combine different generators in a list;
to be read as "or"
- `AsOne` -- the  vector wrapped by `AsOne` is used as one argument value in
an evaluations instead of element-by-element (as scalars)
- `Replicate` -- produces a vector (list) of values generated by a given
generator or expression multiple times repeatedly
- `Anything` -- a convenience wrapper around a few most typical
generators

#### ## "Safety brakes"

Terminators if something goes wrong in tests:

- `stopIfErr`
- `stopIfWarn`
- `stopIfErrOrWarn`

## # Examples

```{r}
library(gentest)
# A trivial example to demonstrate
# the data frame generated by `gentest`
res1 <- gentest(paste,
                x = c('a','b','c'),
                y = 1:2,
                z = list(100.5, 'Zzzzz'),
                sep = '_')
res1
```

```{r}
library(magrittr) # for the pipe operator: %>%
tagNA <- function(x) # Helper function used below
    if (x %>% is.na)
        x %>% gentest:::addClass('NA') else x
# Assume we want to test the function
# `mean(x, na.rm)` paying attention to the
# combinations of arguments when NA is returned:
res2 <- gentest(mean,
                tagNA, # this function is applied to the value returned by `mean`
                x = NumericInRange(-1e6,1e6) %|%
                    String(string_length=3) %|% # max 3 chars to fit the table when printed below
                    (NumericInRange(-1e6,1e6) %>% withNA),
                na.rm = T %|% F)
# Now lets' generate file 'mean.Rd' in sub-directory 'man'
# of the current working directory:
docgen(res2, 'Type-checked mean()')
```

```{r, results='asis', echo=FALSE, warning=FALSE}
genHTML('man/mean.Rd')
```

## # Using your own generator with `docgen`

A generator can be any function. If you want to use the result
of the `gentest` to create automatically the .Rd doc/help file,
it's best (though not mandatory) to tag the return
value of your generator with a human-readable name
through the function `Name`:

```{r}
# Assume you have a generator function that
# generates one of the 3 characters:
# either A or B or C
A_or_B_or_C <- function()
    c('A','B','C') %>%
    # add a human-readable name via Name()
    # if you want to use that generator in
    # `docgen`
    Name('A or B or C')

res3 <- gentest(paste0,
                # `arg1` could be also specified as:
                # 'A' %|% 'B' %|% 'C'
                arg1 = A_or_B_or_C(),
                arg2 = NumericInRange(1,10))
docgen(res3)
```

```{r, results='asis', echo=FALSE, warning=FALSE}
genHTML('man/paste0.Rd')
```

`docgen` generates the documentation only for those
combinations of arguments that do **not** return an error:

```{r}
res4 <- gentest(`+`,
                x = 1:3,
                y = list(101,102,'abc'))
res4
# 'character' as one of the possible classes/types for
# argument `y` is NOT included in the doc file:
docgen(res4)
```

```{r, results='asis', echo=FALSE, warning=FALSE}
genHTML('man/+.Rd')
```

```{r}
# More complex (non-atomic) objects (data frames, lists)
# are also supported:
myfun <- function(df)
    within(df, {
        b <- as.character(a + 1)
        y <- !x
    })
res5 <- gentest(myfun,
                df = Replicate(DataFrame(a = NumericInRange(0, 1e6) %>% withZero,
                                         x = Bool())))
docgen(res5)
```

```{r, results='asis', echo=FALSE, warning=FALSE}
genHTML('man/myfun.Rd')
```
